Deep Blockchains

### 2018-07-05 paper/main.tex
tweaks

\begin{abstract}
We describe how a {\em deep blockchain} architecture can address bottlenecks of a modern single-layer blockchain without sacrificing their core benefits of immutability, security, or trustlessness.  Fundamentally, this is achievable with higher layer blockchains submitting {\em block transactions} containing summaries of the higher layer's block for inclusion in blocks of the lower layer blockchain, with all higher layer blockchains supervenient on lower level blockchain features.   We describe in particular a 3-layer blockchain for provable storage and bandwidth in detail: (I) Layer 1 is MainNet, which stores both (a) registered Layer 3 blockchain roots and (b) Layer 2 Block Merkle roots; (II) Layer 2 is a Plasma Cash chain, storing (a) Plasma tokens redeemable for bandwidth and (b) Layer 3 Block hashes; (III) Layer 3 blockchains are any number of blockchains using storage and bandwidth of Layer 2, e.g those that package NoSQL / SQL transactions for typical database operations;  Layers 2 and 3 utilize a Cloudstore abstraction to store and retrieve blocks and the chunks created by these blocks in Ethereum SWARM and multiple cloud computing providers.  We demonstrate repeated use of Sparse Merkle Trees and show how this construct can be used in our core deep blockchain to provide provable data storage with Deep Merkle Proofs.  We aim to demonstrate implementation results of high-throughput, low-latency layer 3 blockchains resting on an economically sound Layer 2 Plasma Cash blockchains, taken together which are fundamentally capable of scaling for modern web applications.
\end{abstract}

@@

Deep {\em learning} architectures have advanced numerous high-scale applications in every industry in a way that is not about one specific deep learning algorithm -- and instead about an approach that could not be achieved through dogmatic faith in single-layer "neural" networks.   In an analogous way, deep blockchain architectures could have the potential to enable a wide range of high-scale applications in a way that might not be achieved through dogmatic faith in Layer 1 scaling innovations alone.

@@

\noindent Because the retrieval of layer $i+1$ data from layer $i$ can be verified by layer $i+1$ (checking block data: does the block hash match the block content?  is it signed?  does it have a parent hash? etc.; checking chunks: does the hash of the chunk data equal the chunk key), each lower layer node devolves into a dumb storage layer with some failure probability ($p^2_{1}..p^2_{N_2}$ for layer 2 blockchains, $p_1^{1}..p^1_{N_1}$ for layer 1 blockchains -- depending on these failure probabilities, the
Moreover, this deep blockchain uses Ethereum SWARM {\em resource updates} (cite{swarm}), where chunks may be keyed not by the hash of their content but with a {\em resource key}, which can be used for the block data without an index mechanism; all resource updates are signed and so the reader can authenticate.  Ethereum SWARM, because of its use of Kademlia-like protocol, is not naturally as fast as other components in Cloudstore, but the kicks in when all the Layer 2 blockchains fail or when Layer 1 is attacked (51\% attacks, or unknown POS failures).   If other decentralized storage services provided similar provable storage, so long as Layer 3 blockchain does not go Byzantine, only one answer can surface, making for  unstoppable layer 3 blockchains.

@@

Many other Layer 3 blockchains can be constructed using the Layer 2 storage and bandwidth infrastructure: ERC721.  The state of the Layer 3 blockchain is not stored locally but instead kept in Cloudstore with storage and bandwidth costs properly accounted for using the Layer 2 tokens, themselves based on Layer 1.  Layer 3 and Layer 2 nodes are therefore ``light nodes'' in that they can quickly catch up to the latest state by asking the layer 2 and layer 1 blockchains for the most recent finalized block.  This is not possible to do for the Layer 1 blockchain, however.  However, it is possible, and interesting to adapt a Layer 1 blockchain of Ethereum and make it a Layer 3 blockchain.  Computation (Ethereum gas costs) can consume {\em Layer 2 token} balances in state channels along with bandwidth, contract storage can use SMTs mapped to Cloudstore (instead of Patricia Merkle Tries kept in local store) submitted in blocks to the Layer 2 blockchain, and the consensus machinery can be put in a modern sharded Proof-of-Stake framework to achieve high-throughput low-latency ambitions of Ethereum 2.0, with all layer 3 nodes.  The expectation would be that a Layer 3 Ethereum blockchain would have massively lower costs due to rational models of storage and bandwidth.   Other deep blockchain systems can be developed with different computational primitives than the EVM, such as Amazon's Lambda or Apache Hadoop.

>>>>

\begin{abstract}
We describe how a {\em deep blockchain} architecture can address bottlenecks of a modern single-layer blockchain without sacrificing their core benefits of immutability, security, or trustlessness.  Fundamentally, this is achievable with higher layer blockchains submitting {\em block transactions} containing summaries of the higher layer's block for inclusion in blocks of the lower layer blockchain, with all higher layer blockchains supervenient on lower level blockchain features.   We describe in particular a 3-layer blockchain for provable storage and bandwidth in detail: (I) Layer 1 is MainNet, which stores both (a) registered Layer 3 blockchain roots and (b) Layer 2 Block Merkle roots; (II) Layer 2 is a Plasma Cash chain, storing (a) Plasma tokens redeemable for bandwidth and (b) Layer 3 Block hashes; (III) Layer 3 blockchains are any number of blockchains using storage and bandwidth of Layer 2, e.g those that package NoSQL / SQL transactions for typical database operations;  Layers 2 and 3 utilize a Cloudstore abstraction to store and retrieve blocks and the chunks created by these blocks in Ethereum SWARM and multiple cloud computing providers.  We demonstrate repeated use of Sparse Merkle Trees and show how this construct can be used in our core deep blockchain to provide provable data storage with Deep Merkle Proofs.  We aim to demonstrate implementation results of high-throughput, low-latency layer 3 blockchains resting on  economically secure Layer 2 Plasma Cash blockchains, taken together which are fundamentally capable of scaling for modern web applications.
\end{abstract}

@@

Deep {\em learning} architectures have advanced numerous high-scale applications in every industry in a way that is not about one specific deep learning algorithm -- and instead about an approach that could not be achieved through dogmatic faith in single-layer ``neural'' networks.   In an analogous way, deep blockchain architectures could have the potential to enable a wide range of high-scale applications in a way that might not be achieved through dogmatic faith in Layer 1 scaling innovations alone.

@@

\noindent Because the retrieval of layer $i+1$ data from layer $i$ can be verified by layer $i+1$ (checking block data: does the block hash match the block content?  is it signed?  does it have a parent hash? etc.; checking chunks: does the hash of the chunk data equal the chunk key), each lower layer node devolves into a dumb storage layer with some failure or attack probability ($p^2_{1}..p^2_{N_2}$ for layer 2 blockchains, $p_1^{1}..p^1_{N_1}$ for layer 1 blockchains -- depending on this probabilistic model, the cost of attack may be divined.   However, it seems most likely that motivated parties would attack the centralized control behind each layer (c.f. via EIP999, mining pools arewedecentralizedyet.com, governments asking the Cloudstore providers to block Layer 2 operator's accounts) -- in this sense the probabilistic independence in concentrated efforts to attack layer 1 and 2 would be highly suspect.   For this reason, our true faith relies in Ethereum SWARM's {\em resource updates} (\cite{swarm}), where chunks may be keyed not by the hash of their content but with a {\em resource key}, which can be used for the block data without an index mechanism; all resource updates are signed so the reader can authenticate.  Ethereum SWARM, because of its use of Kademlia-like protocol, is not naturally as fast as other components in Cloudstore, but kicks in when all Layer 2 blockchains Cloudstore fail or when Layer 1 itself is attacked (via 51\% attacks, or unknown POS failures).   If other decentralized storage services provided similar provable storage as Ethereum SWARM's resource update, so long as Layer 3 blockchain does not go Byzantine, only one answer can surface, making for unstoppable layer 3 blockchains.

@@

Many other Layer 3 blockchains can be constructed using the Layer 2 storage and bandwidth infrastructure: a chain that represents the evolving state of ERC721 tokens, a chain that represents a cryptocurrency exchange where your money can never be stolen, and so forth.  The state of the Layer 3 blockchain is not stored locally but instead kept in Cloudstore with storage and bandwidth costs properly accounted for using the Layer 2 tokens, themselves based on Layer 1.  Layer 3 and Layer 2 nodes are therefore ``light nodes'' in that they can quickly catch up to the latest state by asking the layer 2 and layer 1 blockchains for the most recent finalized block.  This is not possible to do for the Layer 1 blockchain, however.  However, it is possible, and interesting to adapt a Layer 1 blockchain of Ethereum and make it a Layer 3 blockchain.  Computation (Ethereum gas costs) can consume {\em Layer 2 token} balances in state channels along with bandwidth, contract storage can use SMTs mapped to Cloudstore (instead of Patricia Merkle Tries kept in local store) submitted in blocks to the Layer 2 blockchain, and the consensus machinery can be put in a modern sharded Proof-of-Stake framework to achieve high-throughput low-latency ambitions of Ethereum 2.0, with all layer 3 nodes.  The expectation would be that a Layer 3 Ethereum blockchain would have massively lower costs due to rational models of storage and bandwidth.   Other deep blockchain systems can be developed with different computational primitives than the EVM, such as Amazon's Lambda or Apache Hadoop.
